Wavelet Spatio-Temporal Change Detection on multi-temporal SAR images

### 2021-11-11 tex/paper.wecs/wecs_tgrs.tex
notação no .tex

Let $\mathcal{I}(1),\ldots,\mathcal{I}(m)$ be a set of matrices representing the images of some region of interest. These images may be relative to one SAR channel or a combination of channels; this will be specified when appropriate. Our goal is twofold: to find possible points in time where some relevant changes might have taken place at the region represented in $\mathcal{I}(m)$, $m=1,\ldots,n$, and to find which regions are closely associated to the observed changes along time. We shall address these tasks by analyzing the bidimensional stationary discrete wavelet decomposition of $\mathcal{I}(m)$. Stationary wavelets (also known as non-decimated or redundant wavelets) is a traditional de-noising method that can be efficiently applied to two-dimensional signals such as images \cite{coifman1995translation,atto2012multidate,atto2016wavelet}. After application of this wavelet transform to $\mathcal{I}(m)$ at some appropriate resolution level $J\geq 1$, one of its by-products is a matrix of so called approximation wavelet coefficients $\vX(m)$, a smooth version of $\mathcal{I}(m)$ with the same dimension. The higher $J\in\{1,\ldots,\log_2(k)\}$ is, the smoother $\vX(m)$ gets, where $k$ is the minimum between the numbers of rows and columns of $\mathcal{I}(m)$. Many other aspects can be involved in wavelet analysis of images (e.g., different types of wavelet transform, choice of wavelet basis, thresholding of detail coefficients, etc.), but in the current work we focus on $\vX(m)$, which provides a simple way of performing wavelet smoothing and that gives interesting results; extensions based on further aspects of wavelet transforms are straightforward.

We can then consider further apportioning the total $\mathbb{L}_2$ energy of $\left\{\vX(m)\right\}$ as
\begin{equation} 
\sum_{m=1}^n\|\vX(m)\|^2 = n\|\bar{\mathcal{I}}\|^2+2n\langle\bar{\vX}-\bar{\mathcal{I}},\bar{\mathcal{I}}\rangle+\sum_{m=1}^n\|\mathcal \vX(m)-\bar{\mathcal{I}}\|^2,\label{E:eq_ANOVAlogimage}
\end{equation}
where $\bar{\mathcal{I}}=n^{-1}\sum_{m=1}^n\mathcal{I}(m)$, $\bar{\vX}=n^{-1}\sum_{m=1}^n\vX(m)$ and $\|\vA\|^2=\sum_{i=1}^n\sum_{j=1}^m a_{ij}^2$ for a $n\times m$ matrix $\vA=[a_{ij}]$. The last term in (\ref{E:eq_ANOVAlogimage}) measures deviations in time of $\vX(m)$ compared to an {\it average image} $\bar{\mathcal{I}}$, what motivates us to such deviations to detect change points in time. Since each element (pixel) in $\vX(m)$ also has a corresponding sequence of deviations in time, the relation of each local deviation to the overall measure can also be quantified, what allows us to detect changes in space. Such relation shall be computed in the present work with the Pearson correlation, what shares connections with the idea of feature screening employed in high-dimensional regression, as explained further. Other measures of change in time could be evaluated as well, such as squared differences of consecutive times $\|\mathcal \vX(m)-\vX(m-1)\|^2$, $d(\vX(m),\bar{\vX})$ with $d(\cdot,\cdot)$ denoting some distance measure (e.g., Hellinger, Kullback-Leibler, etc.), among many other possibilities.

Let $X_{k,l}(m)$ and $\bar{\mathcal{I}}_{k,l}$ be the entry $(k,l)$ of $\vX(m)$ and $\bar{\mathcal{I}}$, respectively. Computing $D_{k,l}(m)=(X_{k,l}(m)-\bar{\mathcal{I}}_{k,l})^2$, we denote the matrix of squared mean differences as $\vD(m) = [D_{k.l}(m)]$. We then 
analyze the time series given by
\begin{equation} 
\vd(m)= \sum_{k,l}D_{k,l}(m) = \sum_{k,l}(X_{k,l}(m) - \bar{\mathcal{I}}_{k,l})^2,
\label{E:eq_defdm}
\end{equation}
$m=1,\ldots,n$, which measure the temporal overall variation with respect to $\bar{\mathcal{I}}$.

The time points with highest values of $\vd(m)$ represent the images for which the most expressive changes take place, where changes here are measured through $\mathbb{L}_2$ energy. Define the  $n\times p$ matrix
\begin{equation*}
 \vD=\left(
 \begin{array}{c}
 vec(\vD(1))^T\\
 \vdots\\
 vec(\vD(n))^T\\
 \end{array}
 \right),
\label{eq_defmatrixD}
\end{equation*}
where $vec(\vD(m))$ is the $p\times 1$ vector of wavelet coefficients for time $m$ and $p=\#\{k,l\}$ is total number of locations represented by $\vX(m)$, $m=1,\ldots,n$. Sparsity \cite{johnstone2009statistical} on wavelet coefficients plays a special role here.  We suppose a handful of coefficients drive the changes given by $\vd$, so that the effective dimension of $\vD$ (number of locations where relevant changes occur), say $s_d$, is such that $s_d<<p$. This can be represented as the following linear model
\begin{equation} 
\vd=\vD\vbeta+\vxi
\label{E:sparsemodel_d}
\end{equation}
where $\vbeta$ is sparse, i.e., it has $p-s_d$ null elements, and $\vxi^{(d)}$ is some $n\times 1$ random vector of errors.

In order to identify spatio-temporal changes, we employ the idea of ultra-high dimensional correlation screening \cite{fan2020statistical} as follows. For each local squared mean deviation time series, given by individual elements of $\vD(m)$ across $m=1,\ldots,n$, consider the absolute value of its Pearson correlation with the overall squared mean deviations, given by $\vd$:
\begin{equation*}
R_{k,l}= |\corr\left( \vD_{k,l}, \vd\right)|,
\label{def_Rkld}
\end{equation*}
where $\vD_{k,l}=(D_{k,l}(1),\ldots,D_{k,l}(n))^T$ is the time series of squared mean deviations of wavelet coefficients for the two-dimensional index $\{k,l\}$.

@@

We have a matrix $\vR=[R_{k,l}]$ of correlations of ultra-high dimension. Define the set of 
{\it important} indices for changes in images with respect to $\bar{\mathcal I}$ as 
$\mathcal{M}^{*}=\{(k,l): \textit{changes in } \{\mathcal{I}(m)\}_{m=1}^m \textit{ with respect to}$ $\bar{\mathcal{I}}$ \textit{are affected by changes in location corresponding to index} $(k,l)$ $\textit{in such images}\}$.  This set coincides with the non-zero vectorized one-dimensional indices for the sparse representation of $\vbeta$ in (\ref{E:sparsemodel_d}). 
We build the empirical set of indices flagged as corresponding to change locations by
\begin{equation}
\mathcal{M}_{\tau}=\{(k,l):|R_{k,l}|>\tau\},
\label{E:def_Mtaud}
\end{equation}
where $\tau_d>0$ is a convenient threshold value, function of $n$ and $J$. Under some regularity conditions, the probability that $\mathcal{M}_{\tau}$ contains its target set satisfies
\[
P(\mathcal{M}_{\tau}\supset\mathcal{M}^{*})\rightarrow 1,
\]
as $n\rightarrow\infty$ \cite{fan2020statistical}, where $P$ denotes a probability measure. In other words, the empirical set $\mathcal{M}_{\tau}$ has high probability of detecting the correct change locations in $\mathcal{M}^{*}$ when the number of observations $n$ is large.

Further geometrical motivation for our proposal is given as follows. As defined by (\ref{E:eq_defdm}), we expect $\vd$ to be a vector with some few high values, say $s_d$, and $n-s_d$ smaller values. This segregates the multi-temporal images, since the former time points identify the images in which significant changes occur, while the latter indices identify time points with no major changes. Consider $U>L>0$ such that the $s_d$ highest values of $\vd$ are larger than $U$, and the $n-s_d$ smallest values of $\vd$ are smaller than $L$. We also take $\delta=U-L$. The indices defined by  (\ref{E:def_Mtaud}) are such that   
\[
\frac{<\vD_{k,l},\vd>}{\|\vD_{k,l}\|_2\|\vd\|_2}>\tau,
\]
i.e., such that  $\sum_{m=1}^nD_{k,l}(m)\vd(m)>\tau\|\vD_{k,l}\|_2\|\vd\|_2$. This can be rewritten as
\[
\left|\sum_{m:\vd(m)>U} D_{k,l}(m)\right|-\left|\sum_{m:\vd(m)<L} D_{k,l}(m)\right|>\Delta,
\]
for some arbitrary $\Delta>>0$ (which can be a function of $n$ and $J$). Thence, when we employ correlation screening we select the two-dimensional wavelet indices which have the closest empirical directions to the vector of image temporal changes. Thus we are performing a truly spatio-temporal change detection in a single procedure.

@@

In this section we apply the change detection methods above on simulated data of multi-temporal images. The simulated multi-temporal images ($n=80$) is obtained by repeating a sequence of four images with different types of changes plus a noise. Examples of the first four of these images are shown in Figure \ref{F:EllipsoidChanges}. These images are generated from the sum of two matrices: a signal matrix with ones on entries where ellipses occur and zero elsewhere, and a noise matrix with random variables following a standard Gaussian distribution. The first image, $\mathcal{I}(1)$, presents three elongated ellipses. The second image $\mathcal{I}(2)$ has shorter and larger ellipses added. Smaller ellipses are then added to form $\mathcal{I}(3)$ and $\mathcal{I}(4)$. All the changes made that occur among subsequent images can be seen in Figure \ref{F:Changes_methods_images}(a), which displays a matrix of zeros and ones that correspond, respectively, to locations without and with changes along time. Applying WECS to these images, we obtain a matrix $\vR$ of correlations between deviations of each $\mathcal{I}$ entry with the total squared mean deviation. An example of $\vR$ is presented in Figure \ref{F:Changes_methods_images}(b). For some choice of threshold $\tau$ on absolute values of $\vR$, we obtain a matrix of zeros and ones that can be compared with the total true changes displayed in Figure \ref{F:Changes_methods_images}(a).

@@

Figure \ref{F:Changes_methods_images} illustrates a comparison of different approaches to detect accumulated changes. Panel (b) shows the result of WECS using Daubechies wavelet with two null moments (db2) and $J=2$. Panel (c) presents the result of using aggregated absolute differences, a standard approach where the accumulation of changes are measured by a matrix $\vS=\{\vS_{k,l}\}$ with $S_{kl} = \sum_{m=2}^n	|\mathcal{I}_{k,l}(m) - \mathcal{I}_{k,l}(m-1)|$. Finally, in Panel (d) we can see the result if $\vd(m)$ is performed purely on the spatial domain, using $\mathcal{I}(m)$ instead of $\vX(m)$ in the WECS formulation.

@@

Figure \ref{F:EllipsoidChanges_details} presents the different ROC curves for change detection methods applied to the simulated data as follows. The effects of wavelet bases, level of decomposition, deep-learning feature extraction are shown on the ROC curves. We employ the following wavelet bases: Haar;  Daubechies db2; Daubechies db4; Coiflets coif4; Symlets sym2 ; and Symlets sym4. Panel (c) presents the ROC curves for the proposed method under the aforementioned bases. On all instances $J=2$ is employed. Comparing the ROC curves of all options, we can notice that Daubechies db2 and Symlets sym2 are the best choices.  Panel (b) presents the ROC curves for five different levels of decomposition $J=\{1,2,3,4,5\}$ under the Daubechies db2 basis. Levels $J=2,3$ have a clear better performance, with a slight advantage to $J=2$, since it is uses less levels on the decomposition. The overall performance  of $J=2$ warrants its use for the rest of the comparisons. Panel (d) shows how the proposed method performs in comparison to a deep learning feature extraction from a residual learning network \cite{zhang2017beyond}. WECS is applied with db2 wavelets and $J=2$, whereas the other method applies WECS's idea, the only difference being that $\vX(m)$ is replaced by another smooth version of $\mathcal{I}(m)$ that employs neural-networks trained to compete with wavelet smoothing. We can see that the ROC curves for images treated with a deep-learning method has a performance worse than that obtained with WECS. 
We finally have in Panel (a) the proposed WECS with db2 wavelets and $J=2$ compared to two other non-wavelet methods. The first involves computing $\vd(m)$ without wavelet smoothing, i.e., the squared deviations are computed using $\{\mathcal{I}(m)\}$ instead of $\{\vX(m)\}$, and the classic method of analyzing aggregated absolute differences of $\{\mathcal{I}(m)\}$. The ROC curves in Panel (a) show that the proposed WECS outperforms both the other two methods.

@@

We employed the proposed change detection method on a series of 84 multi-date satellite images. The images were taken on a forest region at the border of Brazil and the French Guiana from 2015-12-26 to 2017-12-3. Each image has two channels (VV and VH) and 1538 by 1556 pixels. We perform a change detection wavelet analyses on the combined image by considering each observed entry as $\mathcal{I}_{k,l}(m) = (\mathrm{VV}_{k,l}(m)^2+\mathrm{VH}_{k,l}(m)^2)^{1/2}$, where $\mathrm{VV}$ and $\mathrm{VH}$ represent the matrices of observations from VV and VH channels, respectively.

@@

\begin{figure}[htp!]
\includegraphics[scale=.3]{../../figs/forest_changes_time} 
\caption{Images $\vX(m)$ from times $m$ correspond to top-3 highest values of $\vd(m)$ and the mean image $\bar{\mathcal{I}}$.}
\label{F:forest_change_times}
\end{figure}

>>>>

Let $\mathcal{I}^{(1)},\ldots,\mathcal{I}^{(n)}$ be a set of matrices representing the images of some region of interest and $n$ instants in time. These images may be relative to one SAR channel or a combination of channels; this will be specified when appropriate. Our goal is twofold: to find possible points in time where some relevant changes might have taken place at the region represented in $\mathcal{I}^{(m)}$, $m=1,\ldots,n$, and to find which regions are closely associated to the observed changes along time. We shall address these tasks by analyzing the bidimensional stationary discrete wavelet decomposition of $\mathcal{I}^{(m)}$. Stationary wavelets (also known as non-decimated or redundant wavelets) is a traditional de-noising method that can be efficiently applied to two-dimensional signals such as images \cite{coifman1995translation,atto2012multidate,atto2016wavelet}. After application of this wavelet transform to $\mathcal{I}^{(m)}$ at some appropriate resolution level $J\geq 1$, one of its by-products is a matrix of so called approximation wavelet coefficients $\vX^{(m)}$, a smooth version of $\mathcal{I}^{(m)}$ with the same dimension. The higher $J\in\{1,\ldots,\log_2(k)\}$ is, the smoother $\vX^{(m)}$ gets, where $k$ is the minimum between the numbers of rows and columns of $\mathcal{I}^{(m)}$. Many other aspects can be involved in wavelet analysis of images (e.g., different types of wavelet transform, choice of wavelet basis, thresholding of detail coefficients, etc.), but in the current work we focus on $\vX^{(m)}$, which provides a simple way of performing wavelet smoothing and that gives interesting results; extensions based on further aspects of wavelet transforms are straightforward.

We can then consider further apportioning the total $\mathbb{L}_2$ energy of $\left\{\vX^{(m)}\right\}$ as
\begin{equation} 
\sum_{m=1}^n\|\vX^{(m)}\|^2 = n\|\bar{\mathcal{I}}\|^2+2n\langle\bar{\vX}-\bar{\mathcal{I}},\bar{\mathcal{I}}\rangle+\sum_{m=1}^n\|\mathcal \vX^{(m)}-\bar{\mathcal{I}}\|^2,\label{E:eq_ANOVAlogimage}
\end{equation}
where $\bar{\mathcal{I}}=n^{-1}\sum_{m=1}^n\mathcal{I}^{(m)}$, $\bar{\vX}=n^{-1}\sum_{m=1}^n\vX^{(m)}$ and $\|\vA\|^2=\sum_{i=1}^n\sum_{j=1}^m a_{ij}^2$ for a $n\times m$ matrix $\vA=[a_{ij}]$. The last term in (\ref{E:eq_ANOVAlogimage}) measures deviations in time of $\vX^{(m)}$ compared to an {\it average image} $\bar{\mathcal{I}}$, what motivates us to such deviations to detect change points in time. Since each element (pixel) in $\vX^{(m)}$ also has a corresponding sequence of deviations in time, the relation of each local deviation to the overall measure can also be quantified, what allows us to detect changes in space. Such relation shall be computed in the present work with the Pearson correlation, what shares connections with the idea of feature screening employed in high-dimensional regression, as explained further. Other measures of change in time could be evaluated as well, such as squared differences of consecutive times $\|\mathcal \vX^{(m)}-\vX^{(m-1)}\|^2$, $d(\vX^{(m)},\bar{\vX})$ with $d(\cdot,\cdot)$ denoting some distance measure (e.g., Hellinger, Kullback-Leibler, etc.), among many other possibilities.

Let $X_{k,l}^{(m)}$ and $\bar{\mathcal{I}}_{k,l}$ be the entry $(k,l)$ of $\vX^{(m)}$ and $\bar{\mathcal{I}}$, respectively. Computing $D_{k,l}^{(m)}=(X_{k,l}^{(m)}-\bar{\mathcal{I}}_{k,l})^2$, we denote the matrix of squared mean differences as $\vD^{(m)} = [D_{k,l}^{(m)}]$. We then 
analyze the time series given by
\begin{equation} 
\vd(m)= \sum_{k,l}D_{k,l}^{(m)} = \sum_{k,l}(X_{k,l}^{(m)} - \bar{\mathcal{I}}_{k,l})^2,
\label{E:eq_defdm}
\end{equation}
$m=1,\ldots,n$, which measure the temporal overall variation with respect to $\bar{\mathcal{I}}$.

The time points with highest values of $\vd(m)$ represent the images for which the most expressive changes take place, where changes here are measured through $\mathbb{L}_2$ energy. Define the  $n\times p$ matrix
\begin{equation*}
 \vD=\left(
 \begin{array}{c}
 vec(\vD(1))^T\\
 \vdots\\
 vec(\vD(n))^T\\
 \end{array}
 \right),
\label{eq_defmatrixD}
\end{equation*}
where $vec(\vD^{(m)})$ is the $p\times 1$ vector of wavelet coefficients for time $m$ and $p=\#\{k,l\}$ is total number of locations represented by $\vX^{(m)}$, $m=1,\ldots,n$. Sparsity \cite{johnstone2009statistical} on wavelet coefficients plays a special role here.  We suppose a handful of coefficients drive the changes given by $\vd$, so that the effective dimension of $\vD$ (number of locations where relevant changes occur), say $s_d$, is such that $s_d<<p$. This can be represented as the following linear model
\begin{equation} 
\vd=\vD\vbeta+\vxi
\label{E:sparsemodel_d}
\end{equation}
where $\vbeta$ is sparse, i.e., it has $p-s_d$ null elements, and $\vxi^{(d)}$ is some $n\times 1$ random vector of errors.

In order to identify spatio-temporal changes, we employ the idea of ultra-high dimensional correlation screening \cite{fan2020statistical} as follows. For each local squared mean deviation time series, given by individual elements of $\vD^{(m)}$ across $m=1,\ldots,n$, consider the absolute value of its Pearson correlation with the overall squared mean deviations, given by $\vd$:
\begin{equation*}
R_{k,l}= |\corr\left( \vD_{k,l}, \vd\right)|,
\label{def_Rkld}
\end{equation*}
where $\vD_{k,l}=(D_{k,l}(1),\ldots,D_{k,l}(n))^T$ is the time series of squared mean deviations of wavelet coefficients for the two-dimensional index $\{k,l\}$.

@@

We have a matrix $\vR=[R_{k,l}]$ of correlations of ultra-high dimension. Define the set of 
{\it important} indices for changes in images with respect to $\bar{\mathcal I}$ as 
$\mathcal{M}^{*}=\{(k,l): \textit{changes in } \{\mathcal{I}^{(m)}\}_{m=1}^m \textit{ with respect to}$ $\bar{\mathcal{I}}$ \textit{are affected by changes in location corresponding to index} $(k,l)$ $\textit{in such images}\}$.  This set coincides with the non-zero vectorized one-dimensional indices for the sparse representation of $\vbeta$ in (\ref{E:sparsemodel_d}). 
We build the empirical set of indices flagged as corresponding to change locations by
\begin{equation}
\mathcal{M}_{\tau}=\{(k,l):|R_{k,l}|>\tau\},
\label{E:def_Mtaud}
\end{equation}
where $\tau_d>0$ is a convenient threshold value, function of $n$ and $J$. Under some regularity conditions, the probability that $\mathcal{M}_{\tau}$ contains its target set satisfies
\[
P(\mathcal{M}_{\tau}\supset\mathcal{M}^{*})\rightarrow 1,
\]
as $n\rightarrow\infty$ \cite{fan2020statistical}, where $P$ denotes a probability measure. In other words, the empirical set $\mathcal{M}_{\tau}$ has high probability of detecting the correct change locations in $\mathcal{M}^{*}$ when the number of observations $n$ is large.

Further geometrical motivation for our proposal is given as follows. As defined by (\ref{E:eq_defdm}), we expect $\vd$ to be a vector with some few high values, say $s_d$, and $n-s_d$ smaller values. This segregates the multi-temporal images, since the former time points identify the images in which significant changes occur, while the latter indices identify time points with no major changes. Consider $U>L>0$ such that the $s_d$ highest values of $\vd$ are larger than $U$, and the $n-s_d$ smallest values of $\vd$ are smaller than $L$. We also take $\delta=U-L$. The indices defined by  (\ref{E:def_Mtaud}) are such that   
\[
\frac{<\vD_{k,l},\vd>}{\|\vD_{k,l}\|_2\|\vd\|_2}>\tau,
\]
i.e., such that  $\sum_{m=1}^nD_{k,l}^{(m)}\vd(m)>\tau\|\vD_{k,l}\|_2\|\vd\|_2$. This can be rewritten as
\[
\left|\sum_{m:\vd(m)>U} D_{k,l}^{(m)}\right|-\left|\sum_{m:\vd(m)<L} D_{k,l}^{(m)}\right|>\Delta,
\]
for some arbitrary $\Delta>>0$ (which can be a function of $n$ and $J$). Thence, when we employ correlation screening we select the two-dimensional wavelet indices which have the closest empirical directions to the vector of image temporal changes. Thus we are performing a truly spatio-temporal change detection in a single procedure.

@@

In this section we apply the change detection methods above on simulated data of multi-temporal images. The simulated multi-temporal images ($n=80$) is obtained by repeating a sequence of four images with different types of changes plus a noise. Examples of the first four of these images are shown in Figure \ref{F:EllipsoidChanges}. These images are generated from the sum of two matrices: a signal matrix with ones on entries where ellipses occur and zero elsewhere, and a noise matrix with random variables following a standard Gaussian distribution. The first image, $\mathcal{I}^{(1)}$, presents three elongated ellipses. The second image $\mathcal{I}^{(2)}$ has shorter and larger ellipses added. Smaller ellipses are then added to form $\mathcal{I}^{(3)}$ and $\mathcal{I}^{(4)}$. All the changes made that occur among subsequent images can be seen in Figure \ref{F:Changes_methods_images}(a), which displays a matrix of zeros and ones that correspond, respectively, to locations without and with changes along time. Applying WECS to these images, we obtain a matrix $\vR$ of correlations between deviations of each $\mathcal{I}$ entry with the total squared mean deviation. An example of $\vR$ is presented in Figure \ref{F:Changes_methods_images}(b). For some choice of threshold $\tau$ on absolute values of $\vR$, we obtain a matrix of zeros and ones that can be compared with the total true changes displayed in Figure \ref{F:Changes_methods_images}(a).

@@

Figure \ref{F:Changes_methods_images} illustrates a comparison of different approaches to detect accumulated changes. Panel (b) shows the result of WECS using Daubechies wavelet with two null moments (db2) and $J=2$. Panel (c) presents the result of using aggregated absolute differences, a standard approach where the accumulation of changes are measured by a matrix $\vS=\{\vS_{k,l}\}$ with $S_{kl} = \sum_{m=2}^n	|\mathcal{I}_{k,l}^{(m)} - \mathcal{I}_{k,l}^{(m-1)}|$. Finally, in Panel (d) we can see the result if $\vd(m)$ is performed purely on the spatial domain, using $\mathcal{I}^{(m)}$ instead of $\vX^{(m)}$ in the WECS formulation.

@@

Figure \ref{F:EllipsoidChanges_details} presents the different ROC curves for change detection methods applied to the simulated data as follows. The effects of wavelet bases, level of decomposition, deep-learning feature extraction are shown on the ROC curves. We employ the following wavelet bases: Haar;  Daubechies db2; Daubechies db4; Coiflets coif4; Symlets sym2 ; and Symlets sym4. Panel (c) presents the ROC curves for the proposed method under the aforementioned bases. On all instances $J=2$ is employed. Comparing the ROC curves of all options, we can notice that Daubechies db2 and Symlets sym2 are the best choices.  Panel (b) presents the ROC curves for five different levels of decomposition $J=\{1,2,3,4,5\}$ under the Daubechies db2 basis. Levels $J=2,3$ have a clear better performance, with a slight advantage to $J=2$, since it is uses less levels on the decomposition. The overall performance  of $J=2$ warrants its use for the rest of the comparisons. Panel (d) shows how the proposed method performs in comparison to a deep learning feature extraction from a residual learning network \cite{zhang2017beyond}. WECS is applied with db2 wavelets and $J=2$, whereas the other method applies WECS's idea, the only difference being that $\vX^{(m)}$ is replaced by another smooth version of $\mathcal{I}^{(m)}$ that employs neural-networks trained to compete with wavelet smoothing. We can see that the ROC curves for images treated with a deep-learning method has a performance worse than that obtained with WECS. 
We finally have in Panel (a) the proposed WECS with db2 wavelets and $J=2$ compared to two other non-wavelet methods. The first involves computing $\vd(m)$ without wavelet smoothing, i.e., the squared deviations are computed using $\{\mathcal{I}^{(m)}\}$ instead of $\{\vX^{(m)}\}$, and the classic method of analyzing aggregated absolute differences of $\{\mathcal{I}^{(m)}\}$. The ROC curves in Panel (a) show that the proposed WECS outperforms both the other two methods.

@@

We employed the proposed change detection method on a series of 84 multi-date satellite images. The images were taken on a forest region at the border of Brazil and the French Guiana from 2015-12-26 to 2017-12-3. Each image has two channels (VV and VH) and 1538 by 1556 pixels. We perform a change detection wavelet analyses on the combined image by considering each observed entry as $\mathcal{I}_{k,l}^{(m)} = ((\mathrm{VV}_{k,l}^{(m)})^2+(\mathrm{VH}_{k,l}^{(m)})^2)^{1/2}$, where $\mathrm{VV}$ and $\mathrm{VH}$ represent the matrices of observations from VV and VH channels, respectively.

@@

\begin{figure}[htp!]
\includegraphics[scale=.3]{../../figs/forest_changes_time} 
\caption{Images $\vX^{(m)}$ from times $m$ correspond to top-3 highest values of $\vd(m)$ and the mean image $\bar{\mathcal{I}}$.}
\label{F:forest_change_times}
\end{figure}
